---
title: "Linear Models workshop (#4) of the R QCBS Workshops Series"
topics: Linear Regression
date: 8/11/2024
link: https://r.qcbs.ca/workshops/r-workshop-04/ 
---
```{r}
# 1. Load libraries
library(car)
library(dplyr)
library(vegan)
library(MASS)
library(effects)
library(e1071)
library(gtsummary)
```

# 2. Load data and explore it
```{r}
# 2.1 Load data
bird <- read.csv("birdsdiet.csv", stringsAsFactors = TRUE)
```


```{r}
# 2.2 Structure of the dataframe
str(bird)
```


```{r}
# 2.3 mean of the response variable
mean(bird$MaxAbund)
```
```{r}
# 2.4 Median of the response variable
median(bird$MaxAbund)
```
```{r}
# 2.5 Variance of the response variable
var(bird$MaxAbund)
```
```{r}
# 2.6 Plot response vs predictor
plot(bird$Mass, bird$MaxAbund)
```
```{r}
# 2.7 Assumptions of the linear models
## 1. Linear relationship between the response and the predictor variables
## 2. Normal distribution of the residuals
## 3. Homoscedasticity
## 4. Independence of residuals
```

# 3. Linear regression in R
## Steps:
### 1. Formulate and run the model
### 2. Check if the conditions are met
### 3. Examine the output of the model - analyze the parameters, plot the model and perfomr significane test on the parameter estimates (if necessary)
```{r}
# 3.1 Formulate the linear model
lm1 <- lm(MaxAbund ~ Mass, data = bird)
```

```{r}
# 3.2 Examine the output
lm1
```
```{r}
# 3.3 Look at the regression table
tbl_regression(lm1)
```
```{r}
# 3.4 Check if the conditions are met
par(mfrow = c(2,2))
plot(lm1)
```
```{r}
# 3.5 Look at the coefficient value
coef(lm1)
```
```{r}
# 3.6 plot the fitted line and the residuals
par(mfrow=c(1,2))
plot(MaxAbund ~ Mass, data = bird)
abline(lm1)
hist(residuals(lm1))
```
```{r}
# 3.7 Test the normality of the residuals
shapiro.test(residuals(lm1))
```
```{r}
# 3.8 Test for skewness in the residuals
skewness(residuals(lm1))
```
## Positive values of the skewness indicates the distribution is left-shifted
## The assumptions of linear models are not met here. How to proceed?
## One solution is to transform the data
```{r}
# 3.9 run regression on the trasformed data
bird$logMaxAbund <- log10(bird$MaxAbund)
bird$logMass <- log10(bird$Mass)
lm2 <- lm(logMaxAbund ~ logMass, data = bird)
```

```{r}
# 3.10 Look at the diagnostic plot
par(mfrow = c(2,2))
plot(lm2)
```


```{r}
# 3.11 Look at regression table
tbl_regression(lm2)
```
```{r}
# 3.12 Plot the linear model observations
par(mfrow = c(1, 2))
plot(logMaxAbund ~ logMass, data = bird)
abline(lm2)
hist(residuals(lm2))
```
```{r}
# 3.13 Look at the summary of the model
summary(lm2)
```

```{r}
# 3.14 Finding a better model
## New hypothesis --> average mass of individuals has an impact on the maximum abundance of terrestrial birds
lm3 <- lm(logMaxAbund ~ logMass, data = bird, subset = !bird$Aquatic)
```

```{r}
# 3.15 Diagnostic plot
par(mfrow = c(2, 2))
plot(lm3)
```
```{r}
# 3.16 Summary\
summary(lm3)
```
```{r}
sessionInfo()
```

